Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-07 Referee1_Response.tex

\noi xxx 
We thank the referee for this suggestion. 
You are right.
There are cases when generating scenarios is expensive in the stochastic optimization setting as well. 
We changed the introduction based on your, the associate editor and the other reviewer's comments. 
We specifically added this as a motivation. 
\medskip

\begin{itemize}
\item[] \textit{However, perhaps more often, sampling the i.i.d. realizations is computationally cheap compared to solving the resulting SAA instances in stochastic optimization. 
In this latter case, we have a choice between simply increasing the number of batches in a non-overlapping method or using the overlapping batch means approach (with some degree of overlap) in order to reduce variance.
Based on the paper's results, perhaps the computational effort required to achieve a level of variance reduction should guide this choice. 
The authors do point to the fact that warm starts in a stochastic
optimization algorithm can be exploited much more easily when using overlapping batch means than when not using overlapping batches. 
Some attempt to explore this tradeoff would be valuable. 
The computational results use sorting (for the news vendor problem), a dynamic programming algorithm (for a stochastic knapsack model), and a decomposition method (for two-stage stochastic linear programs). 
While I am not suggesting that this exploration needs to be made for all three types of algorithms, even exploring it in the case of the sorting algorithm (where some algorithms are well suited for almost-sorted lists) could provide valuable insight.}
\end{itemize}

\noi 
Based on your and the associate editor's comments, we studied computational efficiency of the OMRP variance estimators. 
The results of this study are presented in a new section, \S 5.4 (Computational Efficiency).
%We studied the computational efficiency of the OMRP estimators, based on your and the associate editor's comments, and provided a summary of our results in a new section, Section 5.4 (Computational Efficiency). 
We first approached efficiency by the reciprocal of variance multiplied by the computation time, per the associate editor's suggestion. 
Maximizing computational efficiency defined in this manner provided guidelines on selecting the degree of overlap. 
We then looked at how OMRP compared to increasing the number batches in MRP to reach an equivalent variance level.
We examined computational efficiency theoretically, assuming a functional form of the solution time per batch, and also empirically. 
For empirical evaluation, we used the newsvendor problem with insertion sort, based on your suggestions. 
We found that when effective warm-starting is available, overlapping can lead to more efficient estimators than (nonverlapping) MRP, both when the direct nonoverlapping counterpart is used or when the number of batches in MRP is increased to match the variance of the overlapping counterpart.  
All these results are presented in Section 5.4 (Computational Efficiency). 
Thank you for this suggestion. 
We believe the study of computational efficiency significantly improved the paper. 
(Please also see our response to your minor comment 19 below.)
\medskip

@@

\begin{itemize}
\item[\textbf{M2.}] \textit{The choice of the lower bound that is embedded in equation (8) is interesting. 
(The upper bound in equation (8) is completely natural as the sample mean of the $n_l$ realizations of the function evaluated at $\hat{x}$.) 
It seems that there are also other natural choices that could be made here for the lower bound. 
There is some discussion of this choice in the text that follows equations (7)-(9), but can the authors say more? For example, do the consistency and asymptotic validity results that follow hinge on this choice? Or, do those results also hold for other choices, like forming $\gbb$ as a sample
mean of the underlying $\gb$ values. 
We know that the individual gap estimators in (7) are nonnegative, and if we formed $\gbb$ as a sample mean of $\gb$ values, it would be nonnegative. 
Is the $\gbb$, as specified in (8), nonnegative? 
Have the authors computationally explored other choices here? 
Any insight in this vein would be valuable to the reader in understanding the choice made in equation (8).}
\end{itemize}

\noi
You are right; we did not fully motivate this choice of $\gbb$.  
We added explanations after the definition of the OMRP estimators, motivating this choice and relating it to the theoretical results presented in the paper.\smallskip

\noi 
As you pointed out for the upper bound in (8), this $\gbb$ was chosen for each observation to have equal weight when the sampling problems yield optimal solutions $x_j^* = x^*, \ \forall j$. 
This fact is used in Lemmas 1 and 2, which are then used---directly or indirectly---in proving Theorems 1--3. 
When the alternative $\gbb$, formed as a sample mean of $\gb$ values, is used, 
a subset of the theoretical results can be established (e.g., $\gbb$ converges almost surely to $\mu_{\hat{x}}$). 
However, we had difficulty establishing the same asymptotically lower variances for the new $VG$. 
An alternative is to concurrently adjust the coefficients of the $\dbb$ terms to establish convergence of the new $\gbb$, $VG$, etc.\ to their nonoptimized counterparts $\dbb$, $VD$, etc.
When the coefficients of $\dbb$ are adjusted, however, $\dbb$ and $VD$ are {\it not exactly} the same as the overlapping estimators of simulation output analysis. 
Because our results depend on the asymptotic properties of the overlapping estimators already established in the simulation literature, we did not want to make any changes to $\dbb$. 
It may be still be possible to recover the same results using this alternative  $\dbb$ but we did not pursue it here. 
%We actually performed computations using $\gbb$ formed as a sample mean of $\gb$ values and the results were largely similar. 
Due to the lack of theoretical results explained above, we decided not to include or discuss $\gbb$ formed as a sample mean of $\gb$ values, even though it appears a natural choice. 
Instead, we expanded the text motivating the current choice of $\gbb$. 
\medskip

@@

\begin{itemize}
\item[12.] \textit{ p14, l23: Please add $j = 1$ to the sum.}
\end{itemize}

\noindent 
Added. Thanks for catching this.
\medskip

\begin{itemize}
\item[13.] \textit{p14, l35: The authors have gone to some length to distinguish different types of $A_j$ terms.
However, this equation has an $\mathbb{E}A_j^2$ term that has been collapsed from a sum, but we don't know ``which $j$''. 
Should there be a $\max_j \mathbb{E}A_j^2$ or some other treatment here?}
\end{itemize}

\noindent 
This is because the center batches are identically distributed. 
As a result, $\mathbb{E}A_j^2$ for all the center batches $j=\lceil \bar{\gamma}^{-1} \rceil, \ldots,n_b - \lceil \bar{\gamma}^{-1} \rceil$ are the same. 
To make this clear, we introduced the notation $A_j^c$ for center batches and $A_j^f$ and $A_j^e$ for the fringe batches that appear at the front and end of each set of nonoverlapping batches, respectively. 
We also emphasized that each front fringe batch has an identically distributed pair at the end fringes, and that all the center batches are identically distributed. 
%Thanks for catching this. 
\smallskip

\noi 
{\it Note:} We added brackets to expectations to improve readability as suggested by the other referee.
\medskip

@@

\begin{itemize}
\item[14.] \textit{p15, Theorem 3: A hypothesis in the theorem is that $\hat{x} \neq x^*$, and yet the proof starts with the case that $\hat{x} = x^*$. 
That hypothesis could be removed if the claim had a ``$\geq$'' or this statement could be removed or modified.}
\end{itemize}

\noindent Thank you for pointing this out and for your careful review.  
We removed this statement from the proof of Theorem 3 and added it as a remark to the paragraph after the theorem.
\medskip

### 2014-07-07 AE_Reponse.tex
Moving along
1) Major comment 2 of Ref1
2) Response to Major 1,2 of Ref1
3) Edited response to AE

\noi 
Thank you for this suggestion. 
It is indeed important to study computational efficiency in this context. 
We created a new section, Section 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
When an algorithm has warm-starting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We discussed efficiency for a given functional form of the solution time per batch. 
Depending on the effectiveness of warm-starting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
We illustrated this method on the newsvendor problem using insertion sort. 
Our experiments show that (i) the functional form used provides a good representation, (ii) increased computational efficiency over the nonoverlapping counterpart is indeed observed over a range of overlapping levels, and (iii) this analysis provides a way to select the degree of overlapping. 
We discussed guidelines on selecting the degree of overlapping. 
\medskip

In addition, based on the comments of Referee 1, we studied the efficiency of overlapping compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with lower variances. 
Our results indicate that when warm-starting is effective, overlapping is computationally more efficient than increasing the number of batches of MRP.
Even in cases where warm-starting is less effective, computational efficiency can be observed when about 10\% of observations are overlapped.
We again illustrated this result on the newsvendor problem. 
We believe the paper has improved as a result of the study of computational efficiency. 
Thank you. 
\medskip

@@

\noi xxx 
Based on your other comments and suggestions of the referees, we provided information on computational efficiency gains of overlapping in the context of assessing solution quality in stochastic programming. 
These are detailed in Section 5.4 (Computational Efficiency). 
(Please see our response to the previous comment.)
We also changed Section 1 (Introduction) and Section 6 (Conclusions) to clearly motivate the study and usefulness of the results.

>>>>

\noi 
Thank you for this suggestion. 
It is important to study computational efficiency in this context. 
We created a new section, \S 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
When an algorithm has warm-starting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We discussed efficiency for a given functional form of the solution time per batch. 
Depending on the effectiveness of warm-starting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
\medskip

In addition, based on the comments of Referee 1, we studied efficiency of overlapping compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with lower variances. 
Our results indicate that when warm-starting is effective, overlapping is computationally more efficient.
Even in cases where the effectiveness of warm-starting is quite low, computational efficiency can be observed with a low degree of overlap (e.g., when approximately 10\% of observations are overlapped).\medskip

We illustrated these results on the newsvendor problem using insertion sort. 
Our experiments show that (i) the functional form used provides a good representation, (ii) increased computational efficiency over the nonoverlapping counterpart is indeed observed over a range of overlapping levels, and (iii) this analysis provides a way to select the degree of overlapping by maximizing efficiency. 
Based on our study, we provided guidelines on selecting the degree of overlapping at the end of \S 5.4 (Computational Efficiency). 
\medskip

We believe the paper has improved as a result of the study of computational efficiency. 
Thank you for this suggestion. 
\medskip

@@

\noi 
We established improved computational efficiency on the newsvendor problem.
A study of computational efficiency gains of overlapping in the context of assessing solution quality in stochastic programming is provided in the new \S 5.4 (Computational Efficiency), based on your and the referee's comments. 
(Please see our response to the previous comment.)
\medskip

xxx 
We also changed Section 1 (Introduction) to clearly motivate the study and argue usefulness of the results, paying specific attention to efficiency gains.
Finally, we updated Section 6 (Conclusions) to summarize these new results, again paying attention to computational efficiency.

### 2014-07-07 omrp_lb_rev.tex
Moving along
1) Major comment 2 of Ref1
2) Response to Major 1,2 of Ref1
3) Edited response to AE

>>>>

The motivation behind forming $\gbb_l$ by taking into account $B(i)$ is to give each observation $\xit^i$ equal weight. 
This is because we want to be as close to a standard overall mean estimator (e.g., $\ybb=\frac{1}{n}\sum_{i=1}^{n} y^i$) as possible.
With this definition of $\gbb_l$, when $\xs_j= \xs$ for all $j= 1, 2, \dots, \nbl$, we have $\gbb_l$ equal to its nonoptimized counterpart given in (\ref{eq:dbb}). 
This fact is key in establishing the theoretical results presented in the next section. 
We note, however, that this choice of $\gbb_l$ may not be guaranteed to be non-negative.
In that case, $\max\{\gbb_l,0\}$ can be used in practice. 
In our computational results, however, we did not observe any negative values of $\gbb_l$.
